Identifying Effective Motivational Interviewing Sequences Using Automated Pattern Analysis

### 2018-04-14 jhir/latex-manuscript/manuscript.tex
update manuscript

\caption{Hidden Markov Model Emission Matrices}
\label{tab:emission}
\begin{tabularx}{\textwidth}{lXXXp{0.4cm}XXXXXXXXXXXXXXXXXXX}
\hline\noalign{\smallskip}
State & AF & AR & EA & GIN-FON & GIN-FOP & QEB & QEC-HTP & QEC-MLP & QEF & QE-ST & QO & RC-HTP & RC-MLP & RO & RST & SO & SPT & SS & SUM & HU-PO & HU-PW & LUP \\
\noalign{\smallskip}\hline\noalign{\smallskip}
\rowcolor{lightgraycol}
\multicolumn{23}{l}{\textbf{Successful  Sequences}} \\

@@

\caption{Hidden Markov Model Transition Matrices}
\label{tab:transition}  
\begin{tabularx}{\textwidth}{XXXXXX}
\hline\noalign{\smallskip}
State & High Motivation & High Receptivity & Moderate Receptivity & Low Receptivity & Active Feedback \\
\noalign{\smallskip}\hline\noalign{\smallskip}
\rowcolor{lightgraycol}
\multicolumn{6}{l}{\textbf{Successful  Sequences}} \\
High Motivation & 2\% & 9\% & 24\% & 24\% & 41\% \\
High Receptivity & 7\% & 9\% & 20\% & 25\% & 39\% \\
Moderate Receptivity  & 8\% & 19\% & 24\% & 23\% & 27\% \\
Low Receptivity  & 11\% & 26\% & 28\% & 18\% & 17\% \\
Active Feedback  & 25\% & 34\% & 20\% & 15\% & 6\% \\
\rowcolor{lightgraycol}
\multicolumn{6}{l}{\textbf{Unsuccessful  Sequences}} \\
& Ambivalent & Avoidant & & & \\
Ambivalent & 16\% & 84\% & & & \\
Avoidant & 39\% & 61\% & & & \\
\noalign{\smallskip}\hline
\end{tabularx}
\end{sidewaystable}

@@

Frequent communication patterns in successful and unsuccessful patient-counselor communication sequences are shown in Table~\ref{tab:patterns}. Results from frequent pattern mining analysis indicate that reflections of change talk were the most frequent counselor communication behavior in both successful (36.1\%) and unsuccessful sequences (33.7\%). For each pattern, the statistical significance of difference between pattern occurrence in successful and unsuccessful communications was reported by computing p-value using Pearson's chi-square test. Successful sequences were distinguished from unsuccessful sequences by a higher frequency of counselor questions phrased to elicit change talk (30.8\% versus 17.4\%, p= $<$0.0001), statements emphasizing the patient's decision-making autonomy (28.5\% versus 18.6\%, p=0.0012), questions phrased to elicit commitment language (18.1\% versus 11.6\%, p=0.0118), and reflections of commitment language (20.7\% versus 15.1\%, p= 0.0423). In contrast, unsuccessful sequences were characterized by greater frequency of questions to elicit perceived barriers (14.7\% versus 0\%, p= $<$0.0001), reflections of sustain talk (27.1\% versus 15.8\%, p= $<$0.0001), providing information (28.7\% versus 22.1\%, p=0.0254) and other reflections (11.6\% versus 0\%, p= $<$0.0001). In 14.0\% of the successful sequences, reflections of change talk were paired with a question phrased to elicit change talk; this pattern did not appear in unsuccessful sequences. In contrast, in 10.5\% of the unsuccessful sequences, reflections of change talk were paired with information; this pattern did not appear in successful sequences.

>>>>

\caption{Hidden Markov Model Emission Matrices.}
\label{tab:emission}
\begin{tabularx}{\textwidth}{lXXXp{0.4cm}XXXXXXXXXXXXXXXXXXX}
\hline\noalign{\smallskip}
State & AF & AR & EA & GIN-FON & GIN-FOP & QEB & QEC-HTP & QEC-MLP & QEF & QE-ST & QO & RC-HTP & RC-MLP & RO & RST & SO & SPT & SS & SUM & HU-PO & HU-PW & LUP \\
\noalign{\smallskip}\hline\noalign{\smallskip}
\rowcolor{lightgraycol}
\multicolumn{23}{l}{\textbf{Successful  Sequences}} \\

@@

\caption{Hidden Markov Model Transition Matrices.}
\label{tab:transition}  
\begin{tabularx}{\textwidth}{XXXXXX}
\hline\noalign{\smallskip}
State & High Motivation & High Receptivity & Moderate Receptivity & Low Receptivity & Active Feedback \\
\noalign{\smallskip}\hline\noalign{\smallskip}
\rowcolor{lightgraycol}
\multicolumn{6}{l}{\textbf{Successful  Sequences}} \\
High Motivation & 2\% & 9\% & 24\% & 24\% & 41\% \\
High Receptivity & 7\% & 9\% & 20\% & 25\% & 39\% \\
Moderate Receptivity  & 8\% & 19\% & 24\% & 23\% & 27\% \\
Low Receptivity  & 11\% & 26\% & 28\% & 18\% & 17\% \\
Active Feedback  & 25\% & 34\% & 20\% & 15\% & 6\% \\
\rowcolor{lightgraycol}
\multicolumn{6}{l}{\textbf{Unsuccessful  Sequences}} \\
& Ambivalent & Avoidant & & & \\
Ambivalent & 16\% & 84\% & & & \\
Avoidant & 39\% & 61\% & & & \\
\noalign{\smallskip}\hline
\end{tabularx}
\end{sidewaystable}

@@

Frequent communication patterns in successful and unsuccessful patient-counselor communication sequences are shown in Table~\ref{tab:patterns}. Results from frequent pattern mining analysis indicate that reflections of change talk were the most frequent counselor communication behavior in both successful (36.1\%) and unsuccessful sequences (33.7\%). For each pattern, the statistical significance of difference between pattern occurrence in successful and unsuccessful communications was reported by computing p-value using Pearson's chi-square test. Successful sequences were distinguished from unsuccessful sequences by a higher frequency of counselor questions phrased to elicit change talk (30.8\% versus 17.4\%, p= $<$0.0001), statements emphasizing the patient's decision-making autonomy (28.5\% versus 18.6\%, p=0.0012), questions phrased to elicit commitment language (18.1\% versus 11.6\%, p=0.0118), and reflections of commitment language (20.7\% versus 15.1\%, p=0.0423). In contrast, unsuccessful sequences were characterized by greater frequency of questions to elicit perceived barriers (14.7\% versus 0\%, p= $<$0.0001), reflections of sustain talk (27.1\% versus 15.8\%, p= $<$0.0001), providing information (28.7\% versus 22.1\%, p=0.0254) and other reflections (11.6\% versus 0\%, p= $<$0.0001). In 14.0\% of the successful sequences, reflections of change talk were paired with a question phrased to elicit change talk; this pattern did not appear in unsuccessful sequences. In contrast, in 10.5\% of the unsuccessful sequences, reflections of change talk were paired with information; this pattern did not appear in successful sequences.
